Open Experimental Powder X-ray Diffraction Database

### 2024-04-29 chapters/01_introduction.tex
Added todos in introduction
Added idea for paths towards improving simulation

However due to the scarcity of labeled diffractograms originating from experiment, collecting a labeled dataset
of real experimental powder x-ray diffraction data of sufficient size to train a model
that can be useful in practice is next to impossible and is not the aim of this dataset.
Rather we hope that this dataset can enable machine learning researchers to evaluate how closely their simulations
represent actual experimental data, identify what features are missing and modify their simulations algorithms
accordingly to arrive at a simulation that nearly perfectly mirrors how data appears in reality.\\
One very simple approach which could give a first hint of how closely simulations reflect experiment would be to attempt
to train a classifier on simulated and real diffractograms.
If the classifier performs well that would be an indicator that the simulation doesn't reflect experiment sufficiently well.
Note that the converse doesn't necessarily hold as well.
For a set of given sample properties there is only one corresponding simulated diffractogram, but there may an infinite amount of
variations in how the diffractogram of a sample could look like in experiment for example due varying measurement apperatus, noise
, sample impurities and lattice defects.
In order for a simulated dataset to closely resemble experiment it is critical that the simulatio also represents these
statistical variations.
Determining the nature of these variations is a much a harder task than making individual diffractograms look like they
originated from experiment and cannot be accomplished by simply training a classifier.
Nevertheless, we hope that this dataset can at least provide clues of how to take this aspect into account during
simulation as well.\\

Secondly we would like to also present this dataset as a challenge: The fraction of labeled data can be used to
easily numerically evalute the performance of the model on a broad range of data as it appears in experiment, which can be used
to gauge how well models would perform in practice. \\
But even the unlabeled portion of the dataset can be used to test the performance of models: The least a model
has to do in order to prove useful in practice is to make a guess about the crystal structure that is good enough
for Rietveld refinement to converge toward the true structure. \\
Hence, the performance of the model can be tested by evaluating if the inital guess of the model is sufficient for
Rietveld refinment to refine the initial guess to a structure that produces a diffractogram matching the one observed in experiment.
Finally, by inviting the application of models to data as it appears in experiment we would like to bring attention
to some of the practical difficulties that need to be addressed in applying a model to a wide variety of x-ray diffractograms
collected from different sources: For example the angle and the step size of the diffractograms will differ between sources, 
which is something that any model that aims to perform well on this dataset must be equipped to handle.

% Reference library required and analysis software required, reference library may be incomplete
% software training required, doesnt always converge or not to true minimum,
% sometimes sample is truely unknown,  manual labor required in any case

% -> This is too detailed for this paper, but can use it for thesis
%For one, under normal circumstances, the initial guess of the structure requires manually or computationally consulting
%a reference library of pairs of patterns and structures.
%The reference library is an undesirable overhead that needs to be distributed along with the analysis software and
%maintained to remain up to date and interoperable with the anlysis software.
%The industry standard in terms of such reference libraries is the Powder Diffraction File and costs around 1000\$ per
%year to license from its maintainer, the International Centre for Diffraction Data.
%Another issue is that since the number of chemically viable crystal structures is infinite any reference library, no
%matter how

% Crystallographic open database
% https://www.crystallography.net/cod/

%
%
%Machine learning techniques have emerged as a powerful tool in the toolkit of
%materials scientists. While they are often used to make predictions on
%the properties of materials or find new materials with certain properties, an increasingly interesting domain is the automated analysis of raw experimental measurements
%guided by machine learning\supercite{radovicMachineLearningEnergy2018}.
%
%With the advent of high-throughput experiments, the amount of gathered data is
%vast and the analysis often becomes a bottleneck in the processing pipeline
%\supercite{rahmanianEnablingModularAutonomous2022}. Powder X-ray diffraction
%(XRD) is an important measurement technique used to obtain structural
%information from polycrystalline samples\supercite{harrisContemporaryAdvancesUse2001}.
%The diffractograms are an
%information-dense fingerprint of the structure of the material. However,
%analyzing these diffractograms is not an easy task\supercite{holderTutorialPowderXray2019}.
%Full structure solutions and
%Rietveld refinement take time and require expert knowledge, both about the
%analysis technique and the materials class at hand. This is not feasible in high-throughput experiments on a larger
%scale. Therefore, the question arises whether it is possible to automatically
%analyze powder diffractograms with machine learning models trained on large
%amounts of data, making it possible to run inference almost instantaneously.
%
%During the last few years, there have been several studies tackling this
%objective by applying machine learning models to various tasks concerning the
%analysis of powder diffractograms such as phase
%classification\supercite{leeDeeplearningTechniquePhase2020,maffettoneCrystallographyCompanionAgent2021,schuetzkeEnhancingDeeplearningTraining2021,szymanskiProbabilisticDeepLearning2021,wangRapidIdentificationXray2020},
%phase fraction determination\supercite{leeDatadrivenXRDAnalysis2021}, space
%group
%classification\supercite{parkClassificationCrystalStructure2017,oviedoFastInterpretableClassification2019,zalogaCrystalSymmetryClassification2020,vecseiNeuralNetworkBased2019,suzukiSymmetryPredictionKnowledge2020,chakrabortyDeepCrystalStructure2022},
%machine-learning-guided Rietveld
%refinement\supercite{ozakiAutomatedCrystalStructure2020,fengMethodArtificialIntelligence2019},
%extraction of lattice
%parameters\supercite{dongDeepConvolutionalNeural2021,chitturiAutomatedPredictionLattice2021,chakrabortyDeepCrystalStructure2022,
%habershonPowderDiffractionIndexing2004} and crystallite
%sizes\supercite{dongDeepConvolutionalNeural2021,chakrabortyDeepCrystalStructure2022},
%and also novelty detection based on unsupervised techniques\supercite{bankoDeepLearningVisualization2021}.
%Since an abundant source of experimental diffractograms is hard to come by, most
%applications train their models on simulated diffractograms from the Inorganic
%Crystal Structure Database (ICSD)\supercite{bergerhoff1987}, which contains a total of 272\,260 structures (October 2022).
%
%\citeauthor{leeDeeplearningTechniquePhase2020} used a deep convolutional neural
%network (CNN) trained on a large dataset of multiphase compositions from the
%quaternary \ce{Sr-Li-Al-O} pool to classify present phases in the
%diffractogram\supercite{leeDeeplearningTechniquePhase2020}. In a follow-up
%study, they further showed good results for phase fraction inference in the
%quaternary \ce{Li-La-Zr-O} pool\supercite{leeDatadrivenXRDAnalysis2021}.
%\citeauthor{schuetzkeEnhancingDeeplearningTraining2021} performed phase
%classification on iron ores and cement compounds and used data augmentation with
%respect to lattice parameters, crystallite sizes, and preferred orientation
%\supercite{schuetzkeEnhancingDeeplearningTraining2021}. They showed that
%especially the lattice parameter variations enhance the classification accuracy
%significantly.
%
%Instead of the analysis of phase composition,
%\citeauthor{dongDeepConvolutionalNeural2021} performed regression of scale
%factors, lattice parameters, and crystallite sizes in a five-phase catalytic
%materials system\supercite{dongDeepConvolutionalNeural2021}. In contrast to
%supervised tasks, \citeauthor{bankoDeepLearningVisualization2021} used a
%variational autoencoder to visualize variations in space group, preferred
%orientation, crystallite size, and peak
%shifts\supercite{bankoDeepLearningVisualization2021}.
%\citeauthor{parkClassificationCrystalStructure2017} used a deep CNN to classify
%space groups of single-phase diffractograms, reaching a test accuracy of 81.14\%
%on simulated diffractograms.\supercite{parkClassificationCrystalStructure2017}
%However, as we will show later in this paper, this accuracy is highly overestimated and drops to 56.1\% when test splits are designed in a way to reduce data leakage in non-IID datasets such as the ICSD.
%\citeauthor{vecseiNeuralNetworkBased2019}\supercite{vecseiNeuralNetworkBased2019} developed
%a similar approach and applied their classifier to experimental diffractograms
%from the RRUFF mineral database\supercite{lafuentePowerDatabasesRRUFF2015},
%reaching an experimental test accuracy of 54\%.

% PDF pricing
%https://www.icdd.com/assets/files/2023-Purchase-Options.pdf

>>>>

However due to the scarcity of labeled diffractograms originating from experiment, collecting a labeled dataset
of real experimental powder x-ray diffraction data of sufficient size to train a model
that can be useful in practice is next to impossible and is not the aim of this dataset.
Rather we hope that this dataset can enable machine learning researchers to evaluate how closely their simulations
represent actual experimental data, identify what features are missing and modify their simulations algorithms accordingly to arrive at a simulation that nearly perfectly mirrors how data appears in reality.\\

There are roughly speaking two paths towards improving the simulation: The first is to more accurately model the sample and the processes taking place within it during the simulation of the powder diffraction pattern. This would for example necessitate modelling absorption of the incoming X-ray radiation and its re-emission in the form of characteristic radiation and modeling the effect of stress and motion within the crystal on the diffractogram. \\
An alternative route to obtaining a more accurate simulation is to try to learn these "imperfections" compared to an idealized simulation using machine learning techniques by extracting them from this dataset. \\
As an example of this could be done we can look at results from other domains, where a transfer of core data between different "styles" of representation was attempted:
%\cite{Ganin2015} -> Attempted to

%A first hint of how closely a given simulation reflects experiment could be obtained by
%training a machine learning model, which attempts to classify between simulated and real diffractograms. \\
%If the simulation is very accurate, we would expect the classifier to not have few to no features which it can use to distinguish between simulated and real diffractograms and hence perform poorly. \\
%Conversely, if the classifier performs well and is not difficult to train, that would be an indicator that the simulation doesn't reflect experiment sufficiently well. \\

For a set of given sample properties there is only one corresponding simulated diffractogram, but there may an infinite amount of variations in how the diffractogram of a sample could look like in experiment for example due varying measurement apperatus, noise
, sample impurities and lattice defects.
In order for a simulated dataset to closely resemble experiment it is critical that the simulatio also represents these
statistical variations.
Determining the nature of these variations is a much a harder task than making individual diffractograms look like they
originated from experiment and cannot be accomplished by simply training a classifier.
Nevertheless, we hope that this dataset can at least provide clues of how to take this aspect into account during
simulation as well.\\

Secondly we would like to also present this dataset as a challenge: The fraction of labeled data can be used to easily numerically evalute the performance of the model on a broad range of data as it appears in experiment, which can be used to gauge how well models would perform in practice. \\
But even the unlabeled portion of the dataset can be used to test the performance of models, if the model attempts to predict the entire crystal structure. From the predicted crystal structure, the corresponding diffractogram can be simulated, which can then be compared to the real diffractogram that was used as input for the prediction. While a nearly matching diffractogram doesn't necessarily mean a correctly predicted structure, a non-matching diffractogram certainly means an incorrectly predicted structure. \\

Finally, by inviting the application of models to data as it appears in experiment we would like to bring attention
to some of the practical difficulties that need to be addressed in applying a model to a wide variety of x-ray diffractograms
collected from different sources: For example the angle and the step size of the diffractograms will differ between sources, 
which is something that any model that aims to perform well on this dataset must be equipped to handle.

% TODOS:
% - This paper talks about the importance of sharing raw powder x-ray diffraction data. I'd definitely cite it ->  \cite{Aranda2018}
% - The image diff.png shows the difference between a powder diffractogram calculated from single x-ray diffraction data
% and a real powder diffractogram of the same material. The difference is enormous. I'd include it to highlight
% that simulating powder diffractino patterns from single crystal data only goes so far
%https://www.sciencedirect.com/topics/biochemistry-genetics-and-molecular-biology/rietveld-refinement

%-----------------------------------------------------------%
% Sources on rietveld refinement
